---
title: "shiny.reglog: How to get Google API authorization"
author: "Michał Kosiński"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{shiny.reglog: How to get Google API authorization}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

*shiny.reglog* currently allows for two methods in both data storage and e-mail sending. In both of these areas, one of the method makes use of *gargle*-based packages: *googlesheets4* for data storage and *gmailr* for e-mail sending.

*googlesheets4* API scopes are mid level (reading and writing spreadsheets), so the setup is pretty easy - it needs you to have a Google Account and generate secrets once in interactive session. *gmailr* scope for sending e-mails is high level, so it actually needs you to make some work in Google Cloud Console.

As I know from my own experience, the setup for this authorization may be a little tricky. By writing this vignette I'm trying to get users of *shiny.reglog* to get started quickly with the general settings. For more detailed description I refer to *gargle* sources, mainly:

-   [Non-interactive auth](https://gargle.r-lib.org/articles/non-interactive-auth.html){target="_blank"}
-   [How to use gargle for auth in a client package](https://gargle.r-lib.org/articles/gargle-auth-in-client-package.html#oauth-app){target="_blank"}
-   [How to get your own API credentials](https://gargle.r-lib.org/articles/get-api-credentials.html){target="_blank"}

## Setting up `googlesheets4` auth

To enable authorization for *googlesheets4* in ShinyApp, you need to create a token with correct scopes once, during the interactive session. **auth/spreadsheet** scope don't need the `httr` OAuth app set up (though it is recommended if you are also using *gmailr* for mail sending - details in next section)

```{r, eval = F}
    # Setting up OAuth app isn't mandatory, though it is recommended
    # more details about this in next session

    # app <- httr::oauth_app(
    #   appname = "your app name", # only shows during asking for permission in browser 
    #   key = "Your Client ID", 
    #   secret = "Your Client Secret")

    googlesheets4::gs4_auth(
      email = "gmail used to access the spreadsheets",
      # it is recommended to cache the secret in place where it is easy
      # to find initially. You will need to copy the file to the Shiny Server.
      # After copying it is preferable to hide it for the security
      cache = "path/to/secrets"
      )
```

After running the code once in interactive session, you can copy the secret from **path/to/secrets**, paste it in safe folder on the server and modify the code to show the `gs4_auth()` function where to find it.

Security shouldn't be a problem on [shinyapps.io](https://www.shinyapps.io), but if you are deploying Shiny-Server yourself, remember to secure it - ideally make it available only for the **shiny** user.

## Setting up `gmailr` auth

Scopes for e-mail sending are a bit higher than spreadsheet read & write. You need to use an OAuth app configured in Google Cloud Console. I will try to show in detail how to do it.

-   you need to create a new project in [Google Cloud Console](https://console.cloud.google.com/)

-   enable Gmail API (you can also enable googlesheets API, if you are using **gsheet** method for database) in "APIs & Services" -> "Library"

-   you need to create and configure your app

    -   select "External" user type ("Internal" is only allowed if your account is part of organization)

    -   select correct scopes: "auth/gmail.send" is required for **gmailr** emailing method (you can also select "/auth/spreadsheets" if you are using **googlesheets4** method for databases)

    -   add your gmail email as a "testing user" - if you are going to be using the Google APIs only for your application to get to your APIs, there is no need to Publish your app, it can be "In testing" forever!

-   after configuring your app, you need to create OAuth client

    -   navigate to "Credentials" -> "+ CREATE CREDENTIALS" -> "OAuth Client ID"

    -   select "Desktop app" and input some Name for your app, then click create

    -   you will get a modal dialog with "Your Client ID" (key) and "Your Client Secret" (secret). You will need to provide it to the `httr::oauth_app()` function, as shown below:

```{r, eval = F}
    gmailr::gm_auth_configure(
       key = "Your Client ID", 
       secret = "Your Client Secret"
    )

    gmailr::gm_auth(
      email = "gmail which will be used to send e-mails",
      # it is recommended to cache the secret in place where it is easy
      # to find initially. You will need to copy the file to the Shiny Server.
      # After copying it is preferable to hide it for the security
      cache = "path/to/secrets",
      scopes = "send"
      )
```

After running the code once in interactive session, you can copy the secret from **path/to/secrets**, paste it in safe folder on the server and modify the code to show the `gm_auth()` function where to find it.

Security shouldn't be a problem on [shinyapps.io](https://www.shinyapps.io), but if you are deploying *Shiny-Server* yourself, remember to secure it - ideally make it available only for the **shiny** user.

## Voila!

Congratulations, your ShinyApp should now be able to use the Google API required to access **googlesheets** and/or sending emails through **gmail**!
