---
title: "Creating custom RegLogConnector handlers"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Creating custom RegLogConnector handlers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(shiny.reglog)
```

*shiny.reglog* package requires user to create *dbConnector* and *mailConnector*
for the *RegLogServer* functionality. While creating ShinyApp almost always
you want to use one database and one emailing procedure. For all usage beyond
*RegLogServer* defaults, it would be suboptimal to define new connections. That's
why during development I came to the conclusion that these connectors should
allow for easy extensions with custom functions.

The scope of this vignette is to describe:

- how the *RegLogConnectors* operate
- the default *handlers functions*
- how to write custom *handler function*

## *RegLogConnector* dataflow

Both *dbConnectors* (`RegLogDBIConnector` and `RegLogGsheetConnector`) and
*mailConnectors* (`RegLogEmayiliConnector` and `RegLogGmailrConnector`) inherits
from more general class: `RegLogConnector`. There are three public fields that
are key for the whole dataflow:

- `RegLogConnector$listener()` - *reactiveVal* object that intakes 
*RegLogConnectorMessage* object. *RegLogConnector* listens to any change
in the value of this this object and reacts accordingly.
- `RegLogConnector$handlers` - *named list* of **handler functions**. For every
`type` of received message there should be a specific function appended there.
All of them should return another *RegLogConnectorMessage* object informing
about the result of the function.
- `RegLogConnector$message()` - *reactiveVal* object containing
*RegLogConnectorMessage* that are returned from **handler function**.

### RegLogConnectorMessage object

`RegLogConnector` object reacts upon receiving some kind of 
*RegLogConnectorMessage* object and responds likewise. Its an `S3` class object
that contains four fields:

- `time`: timestamp on which the message was generated
- `type`: character string declaring handler function of the *RegLogConnector*
that should be called when object receive message.
- `data`: list of objects, usually some kind of input or output of the handler
function
- `logcontent`: character string declaring content that should be saved into
*RegLogConnector* logs.

You can create message freely using function of the same name:

```{r preview_of_message}
message <- 
  RegLogConnectorMessage(
    type = "test",
    dataframe = mtcars,
    numbers = runif(10, 0, 100),
    logcontent = paste0("I contain data.frame and random numbers"))

str(message)
```

## Default handler functions for *dbConnectors*

Both *RegLogDBIConnector* and *RegLogGsheetConnector* contain the same default
handler functions. In this vignette I will focus on the messages that are received
by the handler functions and their general usability. To learn about messages
produced by these functions, check "RegLogServer object fields and methods"
vignette and its "Message" section - as all of these messages are finally
exposed in `RegLogServer$message()` public field.

All of these functions aren't exported, as they are used only internally. You can
read the documentation for them though with usual syntax of `?function` in console. 
Documentation is rendered for information how to react with them
by creating *RegLogConnectorMessages* yourself.

### Handler for *login* type message

- *RegLogDBIConnector*: `DBI_login_handler`
- *RegLogGsheetConnector*: `gsheet_login_handler`

These functions are handling querying the database for the specified by the user
of the ShinyApp user ID and password and check if there is a match. Message
structure:

```{r login_message}
login_message <- 
  RegLogConnectorMessage(
    type = "login",
    username = "Whatever",
    password = "&f5*MSYj^niDt=V'3.[dyEX.C/")

str(login_message)
```

### Handlers for *register* type message

- *RegLogDBIConnector*: `DBI_register_handler`
- *RegLogGsheetConnector*: `gsheet_register_handler`

These functions are handling querying the database and checking if the specified
user ID and email for new user aren't already existing in the database. If
there is no conflicts, it will then hash provided password and input
new row. Message structure:

```{r register_message}
register_message <- 
  RegLogConnectorMessage(
    type = "register",
    username = "IAmNewThere",
    email = "something@new.com",
    password = "veryHardP422w0rd!")

str(register_message)
```

### Handlers for *credsEdit* type message

- *RegLogDBIConnector*: `DBI_credsEdit_handler`
- *RegLogGsheetConnector*: `gsheet_credsEdit_handler`

These functions are querying the database to search for the specified username
and password. After confirming user identity, it can update the database row
for this user with any or all of: new username, new email and new password.
Message structure:

```{r credsEdit message}
credsEdit_message <- 
  RegLogConnectorMessage(
    type = "credsEdit",
    password = "&f5*MSYj^niDt=V'3.[dyEX.C/",
    new_username = "Whenever",
    new_email = "edited@email.com",
    new_password = "veryHardP422w0rd!")

str(credsEdit_message)
```

### Handlers for *resetCode_generate* type message

- *RegLogDBIConnector*: `DBI_resetPass_generation_handler`
- *RegLogGsheetConnector*: `gsheet_resetPass_generation_handler`

These functions are querying the database to search for the specified username. 
After confirming that the specified username exists, it generates and inputs
reset code that the user can use to generate new password.
Message structure:

```{r resetPass_generate_message}
resetPass_generate_message <- 
  RegLogConnectorMessage(
    type = "resetPass_generate",
    username = "Whatever")

str(resetPass_generate_message)
```

### Handlers for *resetCode_confirm* type message

- *RegLogDBIConnector*: `DBI_resetPass_confirmation_handler`
- *RegLogGsheetConnector*: `gsheet_resetPass_confirmation_handler`

These functions are querying the database to search for the specified username
and confirming that provided reset code is correct. After confirmation, it
marks the reset code as used and updates password for the user. 
Message structure:

```{r resetPass_confirm_message}
resetPass_confirm_message <-
  RegLogConnectorMessage(
    type = "resetPass_confirm",
    username = "Whatever",
    reset_code = "4265417643",
    password = "veryHardP422w0rd!")

str(resetPass_confirm_message)
```

## Default handler function for *mailConnectors*

All default handlers for *mailConnectors* use the same handler functions:

- *RegLogEmayiliConnector*: `emayili_reglog_mail_handler`
- *RegLogGmailrConnector*: `gmailr_reglog_mail_handler`

They send the email to the specified address using subject and html body of the
email kept in the `mailConnector$mails[[message_type]]` list. They also replace
all of occurences of *?username?*, *?email?*, *?app_name?*, *?app_address?* and
*?reset_code?* with respective values received in the *RegLogConnectorMessage*.

Mail creation is chosen from the `mailConnector$mails` public field on the
basis of the `process` *RegLogConnectorMessage* value.

```{r mail_reglogconnectormessage}
resetPass_mail_message <- 
  RegLogConnectorMessage(
    type = "reglog_mail",
    process = "resetPass",
    username = "Whatever",
    email = "edited@email.com",
    app_name = "RegLog Nice ShinyApp",
    app_address = "https://reglog.nice.com",
    reset_code = "4265417643")

str(resetPass_mail_message)
```

There are also provided handlers to send custom e-mails to the logged user.

- *RegLogEmayiliConnector*: `emayili_custom_mail_handler`
- *RegLogGmailrConnector*: `gmailr_custom_mail_handler`

They are sending email to the specified address parsing it from provided
inside the message *subject* and *body*, providing also an option to send
an *attachment*. No additional parsing is done there and `process` value there
is only informative - it is saved into logs and presented in the
`RegLogServer$mail_message()` field.

```{r custommailmessage}
message_to_send <- RegLogConnectorMessage(
    type = "custom_mail",
    process = "attachment_mail",
    username = "Whatever",
    email = "edited@email.com",
    mail_subject = "Custom message with attachement",
    mail_body = "<p>This is a custom message send from my App</p>
                 <p>It is completely optional, but that kind of message can also
                    contain an attachment!</p>",
    mail_attachement = "files/myplot.png"
  )
```

## How to create custom *handler function*

*Handler function* system of *dbConnectors* and *mailConnectors* allows
for creating custom logic for communicating with them.

For example purposes the custom action that will be described in this vignette
will be the process of saving and reading from *googlesheet* based database
results of simple, 10-item questionnaire: Rosenberg's Self-esteem scale.

### Setup the grounds to store the data

Firstly, we need to create another sheet in the googlesheet that is used
by the *RegLogGsheetConnector* to store our data. Besides the summed score
we will also need timestamp and username to read the most recent row for
the currently logged user.

```{r create_new_sheet, eval=F}
# create new sheet to the googlesheet
googlesheets4::write_sheet(
  ss = gsheet_ss,
  sheet = "SES_results",
  # append 0-row data.frame to create the "schema" for the sheet
  data = data.frame(timestamp = as.character(NA),
                    user_id = as.character(NA),
                    score = as.numeric(NA))[-1,]
)
```

### Create a handler for writing to database

All handler functions need to take as arguments objects `self`, `private` and
`message` and return `RegLogConnectorMessage`.

```{r SES_results_write, eval=F}
write_SES_handler <- function(self, private, message) {
  
  googlesheets4::sheet_append(
    # ID of the connected googlesheet is stored inside private of the
    # RegLogGsheetConnector
    ss = private$gsheet_ss,
    sheet = "SES_results",
    data = data.frame(
      # db_timestamp creates nicely formatted and interpretable by most
      # databases current time
      timestamp = db_timestamp(),
      # user ID and score should be received inside received message
      user_id = message$data$user_id,
      score = message$data$score
      ))
  
  return(RegLogConnectorMessage(type = "write_SES",
                                success = TRUE))
  
}
```

### Create a handler for reading from database

As we have now the writing handler ready, we should create a handler to retrieve
the data in another user session.

```{r SES_results_read, eval=F}
read_SES_handler <- function(self, private, message) {
  
  # read all results
  SES_results <- googlesheets4::read_sheet(
    ss = private$gsheet_ss,
    sheet = "SES_results",
    col_types = "ccn")
  
  # get the lastest result for the current user
  SES_results <- SES_results |>
    dplyr::filter(user_id == message$data$user_id) |>
    dplyr::arrange(dplyr::desc(timestamp)) |>
    dplyr::slice_head()
  
  # return the RegLogConnectorMessage with the score if available
  if (nrow(SES_results) == 1) {
    return(RegLogConnectorMessage(type = "read_SES",
                                  success = TRUE,
                                  score = SES_results$score))
  } else {
    return(RegLogConnectorMessage(type = "read_SES",
                                  success = FALSE))
  }
}
```

### Have everything in motion

I will present there only code for the server logic, containing all needed
elements for appending created *custom handlers* and sending the *RegLogConnectorMessages*
to both write and read data from new sheet.

```{r everything_in_motion, eval=F}
# create and assign RegLogServer object
RegLog <- RegLogServer$new(
  # create googlesheet connector
  dbConector = RegLogGsheetConnector$new(
    # provide correct googlesheet ID
    gsheet_ss = gsheet_ss,
    # provide handlers in a named list. Names will be used to choose on basis
    # of received RegLogConnectorMessage which function to use
    custom_handlers = list(write_SES = write_SES_handler,
                           read_SES = read_SES_handler)
  ),
  # provide some mailConnector with all needed data
  mailConnector = mailConnector
)

# create an event to write the data to the database: there actionButton will
# trigger it
observeEvent(input$write_ses_result, {
  
  # make sure the inputs are provided
  req(input$SES_1, input$SES_2, input$SES_3, input$SES_4, input$SES_5,
      input$SES_6, input$SES_7, input$SES_8, input$SES_9, input$SES_10)
  
  # get the score by summing all raw scores of items
  score <- sum(input$SES_1, input$SES_2, input$SES_3, input$SES_4, input$SES_5,
               input$SES_6, input$SES_7, input$SES_8, input$SES_9, input$SES_10)
  
  # send message to the dbConnector's listener
  RegLog$dbConnector$listener(
    RegLogConnectorMessage(
      # specify correct type - the same as the name of the handler
      type = "write_SES",
      # get required user ID from the RegLog object
      user_id = RegLog$user_id(),
      score = score))
})

# create an event to read the data from the database: eg. another actionButton
observeEvent(input$read_last_ses_result, {
  
  # send correct message to the dbConnector's listener
  RegLog$dbConnector$listener(
    RegLogConnectorMessage(
      type = "read_SES",
      user_id = RegLog$user_id())
  )
})

# assign the retrieved data: eg. to the reactive

SES_result <- reactive(
  # retrieved data will be available in `message()` field of RegLog object
  received_message <- RegLog$message()
  # make sure to only process correct type of message
  req(received_message$type == "read_SES")

  if (!is.null(score)) {
    # get the score if there was any saved in the database
    received_message$data$score
  })
```