---
title: "RegLogServer object fields and methods"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RegLogServer object fields and methods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## About R6 and specifically RegLogServer Class

*R6* is the alternative class system available for **R** users. It is more driven towards
Object Oriented Programming vs Functional Programming that is most known to **R** users
via built-in *S3* class system.

*R6* is ideal for creating more complex systems - it keeps all their elements enclosed
in their own environment. Most of **R** users actually uses *R6* objects on
daily basis: eg. the whole *Shiny* system is built as system of *R6* objects.

Basically, all there is to know about *R6* system in regards to *shiny.reglog*
usage is, that:

- all modules available in *shiny.reglog* are *R6* classes:
  - *RegLogServer*
  - dbConnectors - which are used by every *RegLogServer* to communicate with
  database containing your user data.
     - *RegLogDBIConnector*
     - *RegLogGsheetConnector*
  - mailConnectors - which are used by every *RegLogServer* to send emails to
  users of your ShinyApp.
     - *RegLogEmayiliConnector*
     - *RegLogGmailConnector*
  - vanilla *RegLogConnector*, that you can use to create your own *Connector*.
- R6 objects after initialization have public elements, that can be accessed
  with `$`. They can be either *fields*, holding some values, or *methods*
  containing functions.
  
If you want to get to know more about R6 system as a whole I advise you to check
out [articles created by R6 maintainers](https://r6.r-lib.org/articles/). Below
I will talk more thoroughly about *RegLogServer* class and its public fields and
methods.

There are multiple public elements to *RegLogServer*. To keep them in order, I will
describe them on basis of their usage. In all code chunks I will refer to them
as if they would be called from an object assigned as *RegLogServer*:

```{r setup, include=F}
library(shiny.reglog)
```

```{r RegLogServer_assignement_example, eval = F}
  RegLogServer <- RegLogServer$new(
    dbConnector = dbConnector,
    mailConnector = mailConnector
  )
```

## State of user in current ShinyApp session

Elements described there are the most important things that every creator of 
ShinyApp that incorporates *shiny.reglog* should be using to profit from user
login. All of them are `reactiveVal` objects, so you can `observe` their changes
and to access their current value you need to include them with parentheses.

### Check if the user in current session is logged in

`is_logged` is a simple boolean value to check if the user in current session is 
logged in (`TRUE`) or not (`FALSE`)

```{r is_logged, eval = F}
# observe the reactiveVal change
observeEvent(RegLogServer$is_logged, {
  
  # if contains TRUE: user is currently logged-in
  if (RegLogServer$is_logged()) {
    showModal(modalDialog(title = "You are logged in!"))
    # if contains FALSE: user isn't logged-in
  } else {
    showModal(modalDialog(title = "You are not logged in!"))
  }
})
```

### Current ID of the user

`user_id` field (character string): unique user ID. When the user 
isn't logged it contains generated at the start of the session (or after logout)
*Universally Unique Identifier* (with the use of `uuid::UUIDgenerate()`).
```{r uuid_example}
uuid::UUIDgenerate()
```
After login it contains the username that the user have chosen during registration
procedure.
```{r user_id, eval = F}
RegLogServer$user_id()
```

### Email address of the currently logged user

`user_mail` field (character string): unique user email. When the user is logged in,
it contains the email that is currently associated with their *user_id*. Otherwise,
it contains empty character string (`""`).
```{r user_mail, eval = F}
RegLogServer$user_mail()
```

## Message

`message` field is another `reactiveVal` value that can be used to observe
**any** change of the *RegLogServer* object state. It always contains most recent object 
of *S3* class **RegLogConnectorMessage**, that the *RegLogServer* received from its
*dbConnector* or *mailConnector*, or was specifically generated to communicate
specific status changes.

Creation of custom logic depending on its change isn't at all necessary. It can
be handy especially when for any reason you want to inhibit the default
*modalDialog*s that are called by *RegLogServer* to inform end-user about
consequences of their actions (eg. successful login or unsuccessful, because the
inputed user ID or password is incorrect). 

Every *RegLogConnectorMessage* is specific type of *list* and can contain up to 
four elements:

- **time**: numeric representation of `Sys.time()` when it was generated
- **type**: indicates the type of the process during which the message was generated
- **data**: *list* containing data that was send with this message. Values
contained there can be specific to certain *type* of the message
- **logcontent**: all *RegLogConnectorMessage*s send by and received by the
*RegLogServer* object (and by every *RegLogConnector*) are appended to that
object `log` field. Content there will be saved as a *note* in the logs.

Below I want to characterize all types of *RegLogConnectorMessages* that can be
exposed in the `RegLogServer$message()` field. Besides them there are also messages
that are send by the *RegLogServer* to its *Connectors*. To learn more about these,
read vignette **Creating custom RegLogConnector handlers**.

### Type: **ping**

The object that will be available in the 