---
title: "RegLogServer object fields and methods"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RegLogServer object fields and methods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## About R6 and specifically RegLogServer Class

*R6* is the alternative class system available for **R** users. It is more driven towards
Object Oriented Programming vs Functional Programming that is most known to **R** users
via built-in *S3* class system.

*R6* is ideal for creating more complex systems - it keeps all their elements enclosed
in their own environment. Most of **R** users actually uses *R6* objects on
daily basis: eg. the whole *Shiny* system is built as system of *R6* objects.

Basically, all there is to know about *R6* system in regards to *shiny.reglog*
usage is, that:

- all modules available in *shiny.reglog* are *R6* classes:
  - *RegLogServer*
  - dbConnectors - which are used by every *RegLogServer* to communicate with
  database containing your user data.
     - *RegLogDBIConnector*
     - *RegLogGsheetConnector*
  - mailConnectors - which are used by every *RegLogServer* to send emails to
  users of your ShinyApp.
     - *RegLogEmayiliConnector*
     - *RegLogGmailConnector*
  - vanilla *RegLogConnector*, that you can use to create your own *Connector*.
- R6 objects after initialization have public elements, that can be accessed
  with `$`. They can be either *fields*, holding some values, or *methods*
  containing functions.
  
If you want to get to know more about R6 system as a whole I advise you to check
out [articles created by R6 maintainers](https://r6.r-lib.org/articles/). Below
I will talk more thoroughly about *RegLogServer* class and its public fields and
methods.

There are multiple public elements to *RegLogServer*. To keep them in order, I will
describe them on basis of their usage. In all code chunks I will refer to them
as if they would be called from an object assigned as *RegLogServer*:

```{r setup, include=F}
library(shiny.reglog)
```

```{r RegLogServer_assignement_example, eval = F}
  RegLogServer <- RegLogServer$new(
    dbConnector = dbConnector,
    mailConnector = mailConnector)
```

## State of user in current ShinyApp session

Elements described there are the most important things that every creator of 
ShinyApp that incorporates *shiny.reglog* should be using to profit from user
login. All of them are `reactiveVal` objects, so you can `observe` their changes
and to access their current value you need to include them with parentheses.

### Check if the user in current session is logged in

`is_logged` is a simple boolean value to check if the user in current session is 
logged in (`TRUE`) or not (`FALSE`)

```{r is_logged, eval = F}
# observe the reactiveVal change
observeEvent(RegLogServer$is_logged, {
  
  # if contains TRUE: user is currently logged-in
  if (RegLogServer$is_logged()) {
    showModal(modalDialog(title = "You are logged in!"))
    # if contains FALSE: user isn't logged-in
  } else {
    showModal(modalDialog(title = "You are not logged in!"))
  }
})
```

### Current ID of the user

`user_id` field (character string): unique user ID. When the user 
isn't logged it contains generated at the start of the session (or after logout)
*Universally Unique Identifier* (with the use of `uuid::UUIDgenerate()`).
```{r uuid_example}
uuid::UUIDgenerate()
```
After login it contains the username that the user have chosen during registration
procedure.
```{r user_id, eval = F}
RegLogServer$user_id()
```

### Email address of the currently logged user

`user_mail` field (character string): unique user email. When the user is logged in,
it contains the email that is currently associated with their *user_id*. Otherwise,
it contains empty character string (`""`).
```{r user_mail, eval = F}
RegLogServer$user_mail()
```

## User logout

You can insert an event to logout user by calling public method.

```{r user_logout, eval = F}
# if you create an actionButton "logout_button" in your UI, you can observe
# its action and logout after press:
observeEvent(input$logout_button, {
  RegLogServer$logout()
})
```

## UI object lists

There are dedicated functions for creating UI for different functionalities of the
*RegLog* system: `RegLog_login_UI`, `RegLog_register_UI`, `RegLog_credsEdit_UI`
and `RegLog_resetPass_UI`. They put whole default `tagLists` into the UI, though.

It is expected that there will be users who would want more freedom with how
their UI should look like. That's why there are dedicated public fields containing
named lists of UI elements for all of these functionalities. You can use these
to create your own, custom `renderUI` output.

- `UI_list_login` containing elements of login procedure tags
- `UI_list_register` containing elements of register procedure tags
- `UI_list_credsEdit` containing elements of credentials edit procedure tags
- `UI_list_resetPass` containing elements of reset password procedure tags

## Message

`message` field is another `reactiveVal` value that can be used to observe
**any** change of the *RegLogServer* object state. It always contains most recent object 
of *S3* class **RegLogConnectorMessage**, that the *RegLogServer* received from its
*dbConnector* or was specifically generated to communicate specific status changes.

>Messages received from mailConnector aren't exposed in this field. Instead, they
are exposed in `RegLogServer$mail_message()` field. They are separated because
they don't idicate a change in state of RegLog itself - instead, they just
show the feedback from mail sending service.

Creation of custom logic depending on its change isn't at all necessary. It can
be handy especially when for any reason you want to inhibit the default
*modalDialog*s that are called by *RegLogServer* to inform end-user about
consequences of their actions (eg. successful login or unsuccessful, because the
inputed user ID or password is incorrect). 

Every *RegLogConnectorMessage* is specific type of *list* and can contain up to 
four elements:

- **time**: numeric representation of `Sys.time()` when it was generated
- **type**: indicates the type of the process during which the message was generated
- **data**: *list* containing data that was send with this message. Values
contained there can be specific to certain *type* of the message
- **logcontent**: all *RegLogConnectorMessage*s send by and received by the
*RegLogServer* object (and by every *RegLogConnector*) are appended to that
object `log` field. Content there will be saved as a *note* in the logs.

Below I want to characterize all types of *RegLogConnectorMessages* that can be
exposed in the `RegLogServer$message()` field. Besides them there are also messages
that are send by the *RegLogServer* to its *Connectors*. To learn more about these,
read vignette **Creating custom RegLogConnector handlers**.

>Conditions for every default modalDialog are written in the order they should be
checked in for best results. 
>Conditions written like `value == FALSE` can be checked like that - the *message*
of given type always contains this object. Conditions written like `isFALSE(value)` 
means that the specified *value* in the *message* can be `NULL`.

### Type: **ping**

Type of the message that is produced by all classes inheriting from *RegLogConnector*
upon their initialization, making it also the very first available in `RegLogServer$message()`
field. Received message of this type contains:

- *data*: response_time: time between receiving the **ping** message and sending
this message back
- *logcontent*: for message upon initialization it will hold value "init".

### Type: **login_front**

Type of the message generated by *RegLogServer* object itself. Contains
information about invalidity of values provided during login procedure
by user. It contains:

- **data**
  - success: boolean - `FALSE`
  - input_provided: boolean - `FALSE`
  
This message type is binded with default *modalDialogs*:

- login_noInput (`input_provided == FALSE`)

### Type: **login**

Type of the message received from the database connectors with responses about
login procedure. It contains:

- **data**:
  - success: boolean - `TRUE` if the login was successful
  - username: boolean - `TRUE` if the username exists in the database
  - password: boolean - `TRUE` if the password provided match
  - *user_id*: character - contains user_id presented afterwards in the
  `RegLogServer$user_id()`. Only if `success == TRUE`
  - *user_mail*: character - contains user_mail presented afterwards in the
  `RegLogServer$user_mail()`. Only if `success == TRUE`
- **logcontent**:
  - brief information with username provided by the user and if the login
  was successful.

This message type is binded with default *modalDialogs*:

- login_badId (`data$success == FALSE && data$username == FALSE`)
- login_badPass (`data$success == FALSE && data$password == FALSE`)
- login_success (`data$success == TRUE`)


### Type: **register_front**

Type of the message generated by *RegLogServer* object itself. Contains
information about invalidity of values provided during register procedure
by user. It contains:

- **data**:
  - success: boolean - `FALSE`
  - input_provided: boolean - `TRUE` if all needed inputs were provided
  - valid_id: boolean - `TRUE` if provided user ID was valid
  - valid_email: boolean - `TRUE` if provided email was valid
  - valid_pass: boolean - `TRUE` if provided password was valid
  - identical_pass: boolean - `FALSE` (it is the last condition checked)
  
This message type is binded with default *modalDialogs*:

- register_noInput (`data$input_provided == FALSE`)
- register_nonValidId (`isFALSE(data$valid_id)`)
- register_nonValidEmail (`isFALSE(data$valid_email)`)
- register_nonValidPass (`isFALSE(data$valid_pass)`)
- register_notIndenticalPass (`isFALSE(data$identical_pass)`)

### Type: **register**

Type of the message received from the database connector with responses about
register procedure. It contains:

- **data**:
  - success: boolean - `TRUE` if the register was successful
  - username: boolean - `TRUE` if there were no conflicts with existing 
  usernames in the database
  - email: boolean - `TRUE` if there were no conflicts with existing
  e-mails in the database
  - *user_id*: character - contains user ID with which new user have been
  registered. Only if `success == TRUE`
  - *user_mail*: character - contains e-mail with which new user have
  been registered. Only if `success == TRUE`
- **logcontent**:
  - registered user ID and mail if the register was successful or value
  of the element which caused conflict.
  
This message type is binded with default *modalDialogs*:

- register_existingId (`data$success == FALSE && data$username == FALSE`)
- register_existingEmail (`data$success == FALSE && data$email == FALSE`)
- register_success (`data$success == TRUE`)

### Type: **credsEdit_front**

Type of the message generated by *RegLogServer* object itself. Contains
information about invalidity of values provided during credentials change procedure
by user. It contains:

- **data**:
  - success: boolean - `FALSE` (generated only when something is invalid)
  - input_provided: boolean - `TRUE` if required inputs are not empty
  - valid_id: boolean - `TRUE` if provided user ID was valid
  - valid_email: boolean - `TRUE` if provided email was valid
  - valid_pass: boolean - `TRUE` if provided password was valid
  - identical_pass: boolean - `FALSE` (it is the last condition checked)
  
This message type is binded with default *modalDialogs*:

- credsEdit_noInput_pass (`data$input_provided == FALSE` && change == "pass"`)
- credsEdit_noInput_other (`data$input_provided == FALSE` && change == "other"`)
- credsEdit_nonValidId (`isFALSE(data$valid_id)`)
- credsEdit_nonValidEmail (`isFALSE(data$valid_email)`)
- credsEdit_nonValidPass (`isFALSE(data$valid_pass)`)
- credsEdit_notIdenticalPass (`isFALSE(data$identical_pass)`)

### Type: **credsEdit**

Type of the message received from the database connector with responses about
credentials edit procedure. It contains:

- **data**:
  - success: boolean - `TRUE` if credentials edit was successful
  - username: boolean - `TRUE` if provided username was found
  - password: boolean - `TRUE` if provided password was correct
  - *new_username*: boolean - `TRUE` if provided new ID weren't in
  database (no conflicts). Only if there were some conflict.
  - *new_mail*: boolean - `TRUE` if provided new e-mail weren't in
  database (no conflicts). Only if there were some conflict.
  - *new_user_id*: character - contains new user ID. Only if user ID was changed.
  - *new_user_mail*: character - contains new user e-mail. Only if user mail
  was changed.
  - *new_user_pass*: boolean - `TRUE`. Only if user password was changed.
- **logcontent**:
  - brief information about changes.
  
This message type is binded with default *modalDialogs*:

- credsEdit_badId (`data$success == FALSE && data$username == FALSE`)
- credsEdit_badPass (`data$success == FALSE && data$password == FALSE`)
- credsEdit_existingId (`data$success == FALSE && isFALSE(data$username)`)
- credsEdit_existingEmail (`data$success == FALSE && isFALSE(data$email)`)
- credsEdit_success (`data$success == TRUE`)

### Type: **resetPass_front**

Type of the message generated by *RegLogServer* object itself. Contains
information about invalidity of values provided during password reset procedure
by user. It contains:

- **data**
  - success: boolean - `FALSE` (message generated only when something is invalid)
  - step: character - either `generate` or `confirm`, depending on the step of the process
  - input_provided: boolean - `TRUE` if the inputs required for specific step were provided
  - valid_pass: boolean - `TRUE` if provided password was valid
  - identical_pass: boolean - `FALSE` (it is the last condition checked)
  
This message type is binded with default *modalDialogs*:

- resetPass_noInput_generate (`data$step == "generate" && isFALSE(data$input_provided)`)
- resetPass_noInput_confirm (`data$step == "confirm" && isFALSE(data$input_provided)`)
- resetPass_nonValidPass (`isFALSE(data$valid_pass)`)
- resetPass_notIdenticalPass (`isFALSE(data$identical_pass)`)
  
### Type:  **resetPass_generate**

Type of the message received from the database connector with responses about
password reset procedure - code generation step. It contains:

- **data**:
  - success: boolean - `TRUE` if reset code was created successfully
  - username: boolean - `TRUE` if the username was found in the database
  - email: character - email to which the reset code will be sent
- **logcontent**:
  - information about user for whom the reset code was generated and to which
  email it will be sent.
  
This message type is binded with default *modalDialog*:

- resetPass_badId (`data$username == FALSE`)
- resetPass_codeGenerated (`data$success == TRUE`)
  
### Type: **resetPass_confirm** 

Type of the message received from the database connector with responses about
password reset procedure - code confirmation step. It contains:

- **data**
  - success: boolean - `TRUE` if the password was changed successfully
  - username: boolean - `TRUE` if the username was found in the database
  - code_valid: boolean - `TRUE` if the reset code was valid (both matched and not expired)
- **logcontent**
  - username for which the password was changed.
  
This message type is binded with default *modalDialogs*:

- resetPass_badId (`data$username == FALSE`)
- resetPass_nonValidCode (`isFALSE(data$code_valid)`)
- resetPass_success (`data$success == TRUE`)

### Type: **logout**

Type of the message generated by *RegLogServer* itself, indicating that the
user have been logged out. It contains:

- **data**
  - success: boolean - `TRUE` if the user was logged in and have been logged out
- **logcontent**
  - username of the user that had logged out
  
This message type is binded with default *modalDialog*:

- logout_notLogIn (`data$success == FALSE`)
- logout_success (`data$success == TRUE`)

## Mail message

`mail_message` field contains last *RegLogConnectorMessage* received from
*mailConnector*. It only provides information about last process of sending
email, which don't signalize the change of the data or state of current login
session, so it is separated from much more important `message` field.

### Type: **reglog_mail**

This type of message is received after default e-mail sending from all RegLog
processess. It contains:

- **data**
  - process: character - name of the process for which e-mail was sent. Either
  'register', 'credsEdit' or 'resetPass'
  - success: boolean - `TRUE` if the email was sent sucessfully
- **logcontent**
  - username and address to which email was sent, and the name of the process that
  triggered it. If `success == FALSE` it will also contain error message.
  
### Type: **custom_mail**

This type of message isn't received during regular *RegLog* run. It is a response
to message that you can use to sent a custom e-mail. It contains the same
responses as **reglog_mail** message.

To send a custom email using *mailConnector*, pass to its *listener* a 
*RegLogConnectorMessage* of type *custom_mail*, as in example below:

```{r eval=F}
# you can observe some kind of event to trigger the send
observeEvent(input$send_custom_email, {
  
  # as the username and email will be acquired from RegLogServer,
  # it is best to make sure that the user is logged-in
  req(RegLogServer$is_logged())
  
  message_to_send <- RegLogConnectorMessage(
    type = "custom_mail",
    # name your process in some way - it will be tracked by the app and saved
    # into logs
    process = "custom mail with attachment!",
    # username and email can be gotten from RegLogServer
    username = RegLogServer$user_id(),
    email = RegLogServer$user_mail(),
    # we can specify the subject and body of message ourselves
    mail_subject = "Custom message with attachement",
    # it's best for the body to contain html code
    mail_body = "<p>This is a custom message send from my App</p>
                 <p>It is completely optional, but that kind of message can also
                    contain an attachment!</p>",
    # optionally: attachment
    mail_attachement = "files/myplot.png"
  )
})
```

## Collected logs during the App session

During the lifespan of the session, *RegLogServer* object sends, shows and receives
many different *RegLogConnectorMessages*. By default, all messages **send** to 
its *dbConnector* and *mailConnector* and **received** back are saved into the
`RegLogServer$logs` field, into separate lists per direction. 
There are also some messages that are only **shown** by the object (namely all 
messages with "_front" suffix and message of type **logout**)

They contain following information:

- **time** when the message was generated
- **session** in which the message was generated (taken from `session$token`)
- **type** of the message
- **note** in which the `logcontent` of the message is saved

### Collecting all logs

You can use public method to get whole content of the logs collected up to this
point in time in the form of *data.frame*.

```{r, eval = F}
logs_df <- RegLogServer$get_logs()
```

## Configuring RegLogServer

As the whole *RegLog* system is created with elasticity in mind, there are also
many options for customizing the behaviour of *RegLogServer* object.

### Customizing messages and texts 

During *RegLogServer* object initialization you can specify some arguments:

- `app_name` - name of your application that will show up in default emails
sent to your users. Defaults to the name of root folder of your application.
- `app_address` - URL to your application that will show up in default emails
sent to your users. Defaults to the URL that the App is accessed on the client
side (actually: `NULL`, as the address is got on session start)
- `lang` - language in which the default texts are presented. Defaults to "en"
for English. You can also specify "i18" for the RegLog to show only the content
identifiers for ease of building your own translations using external tools.
- `custom_txts` - defaults to NULL. Can be a named list of character vectors that
should be used in place of defined text for language of your choosing (to 
check them out, use `RegLog_texts()` function).

    You can define custom texts during initialization of *RegLogServer*:
    ```{r, eval = F}
        # initialize new RegLogServer object
    RegLog <- RegLogServer$new(
      # assign created dbConnector and mailConnector
      dbConnector = dbConnector,
      mailConnector = mailConnector,
      # replace default title and body of `login_success` modalDialog
      custom_txts = list(
        login_success_t = "Welcome!",
        login_success_b = "Hi! It's nice to see you logged into our ShinyApp!"
      )
    )
    ```
    
    
- `use_modals` - defaults to TRUE. If specified as `FALSE`, no default modals
will be shown. You can also inhibit only specific modal dialogs by providing
named list of FALSE values. Eg:

    ```{r, eval = F}
    # initialize new RegLogServer object
    RegLog <- RegLogServer$new(
      # assign created dbConnector and mailConnector
      dbConnector = dbConnector,
      mailConnector = mailConnector,
      # inhibit default 'login_success' and 'register_success' modals
      use_modals = list(
        login_success = FALSE,
        register_success = FALSE
      )
    )
    ```

### Options related to log saving

There are two options that can be defined in the script that relates to the log
collection: *RegLogServer.logs* that defines saving logs to `RegLogServer$logs`
and *RegLogServer.logs_to_database* that defines saving the logs into table
or spreadsheet created for log-recording purposes. 

> To save logs into database, remember to firstly create a table or spreadsheet.
During tables creation with either `DBI_tables_create` or `gsheet_tables_create`
specify argument `use_log = T`.

To define options just insert in the beginning of your script:

```{r, eval = F}
# examples with default values
options("RegLogServer.logs" = 1)
options("RegLogServer.logs_to_database" = 0)
```

Possible values:

- 0: no message is saved into logs/database
- 1: messages **send** and **received** are saved
- 2: all messages are saved

> It isn't recommended to save any logs into database when using *RegLogGsheetConnector* 
as your database driver. All actions made on the *googlesheets* database are taking
significant time. Recommended only during testing/debugging runs.

### Options related to e-mail sending

By default, *RegLogServer* will trigger an e-mail send to users of your ShinyApp
after they are registered, when they attempt reset password procedure (to provide
them with their reset code) and after changing their user ID or e-mail during the
credentials edit procedure.

As the e-mail sending is mandatory during reset password procedure, there is no
option to inhibit it, but the other messages are only recommended and are
enriching the user experience. If you are concerned with the amount of e-mails
that can be sent from your app, you can inhibit these messages using following
options:

- `options("RegLogServer.register_mail" = FALSE)`
- `options("RegLogServer.credsEdit_mail" = FALSE)`

To inhibit all e-mail sending (eg. if the app is under development, you are 
debugging it or only playing with the package currently), you can pass an
`RegLogConnector` object to the `mailConnector` argument during initialization
of the *RegLogServer*:

```{r eval = F}
RegLog <- RegLogServer$new(
  dbConnector = somedbConnector,
  mailConnector = RegLogConnector$new()
)
```
